### 2016-04-30 writing/latex/coincidence.tex
tidy up a bit

\chapter[Coincidence Classification]{Coincidences amongst families of mesh patterns and classical patterns}\label{chap:coincs}
One interesting question to ask about permutation patterns considers when a
pattern may be avoided by, or contained in, arbitrary permutations. Two
patterns \(\pi\) and \(\sigma\) are said to be \emph{coincident} if the set
of permutations that avoid \(\pi\) is the same as the set of permutations that
avoid \(\sigma\) \ie \(\av{\pi} = \av{\sigma}\). This extends to sets of patterns as well as single patterns.

We consider the avoidance sets, \(\av{p,q}\) where \(p\) is a classical pattern
of length \(3\) and \(q\) is a mesh pattern of length \(2\) in order to
establish some rules about when these two sets give the same avoidance set.
We fix \(p\) in order to define the equivalences and say that \(p\) is the
\emph{dominating pattern}.

We first define some operations on mesh patterns.
\begin{definition}
Given a mesh pattern \(p\), let \(\addp[D]{p}{(a,b)}\) be the operation that returns
a mesh pattern equivalent to placing a point in the center of box \((a,b)\), where \((a,b)\) is not the
mesh set of \(p\) in
the figure of \(p\), with shading defined by \(D\subseteq\{N,E,S,W\}\).
The set \(D\) defines the shading by indicating that the boxes in the cardinal
directions in \(D\) next to the point are shaded in the resulting pattern.
\end{definition}
Since there is no ambiguity we let \(\addpe[D]\) be equivalent to \(\addp[D]{\varepsilon}{(0,0)}\).

\begin{example}
    The result of adding a single point to the empty permutation for each cardinal direction.
 \begin{align*}
     \addpe[\{N\}] = \textpattern{}{1}{0/1,1/1} &\qquad\addpe[\{E\}] =\textpattern{}{1}{1/1,1/0} \qquad \\
     \addpe[\{S\}] = \textpattern{}{1}{0/0,1/0} &\qquad\addpe[\{W\}] =\textpattern{}{1}{0/0,0/1} \qquad
\end{align*}
    A more complex example for \texttt{add\_point}
\begin{equation*}
\addp[\{N,E\}]{\pattern{}{2,1,3}{0/3}}{(2,3)} = \pattern{}{2,1,4,3}{0/3,0/4,2/4,3/4,3/3}
\end{equation*}
\end{example}
Formally, given a mesh pattern \(p = (\pi,R)\) \(\addp[D]{p}{(a,b)}\) gives a mesh
pattern \(p^\prime = (\pi^\prime,R^\prime)\) defined by
\begin{equation*}
    \pi^\prime(i) = \begin{cases}
    \pi(i) & \text{if } i\neq a+1 \text{ and } \pi(i)<b\\
    \pi(i)+1 & \text{if } i\neq a+1 \text{ and } \pi(i)>b\\
    b+1 & \text{if } i = a+1
\end{cases}
\end{equation*}
and
\begin{equation*}
        R^\prime = \bigcup_{(i,j)\in R}r((i,j))
\end{equation*}
Where \(r((i,j))\) is defined by
\begin{equation*}
    r((i,j)) =
    \begin{cases}
        \{(i,j)\} & \text{if } i<a, j<b\\
        \{(i,j),(i,j+1)\} & \text{if } i<a, j=b\\
        \{(i,j+1)\} & \text{if } i<a, j>b\\
        \{(i,j),(i+1,j)\} & \text{if } i=a, j<b\\
        \{(i,j+1),(i+1,j+1)\} & \text{if } i=a, j>b\\
        \{(i+1,j)\} & \text{if } i>a, j<b\\
        \{(i+1,j),(i+1,j+1)\} & \text{if } i>a, j=b\\
        \{(i+1,j+1)\} &\text{if } i>a, j>b\\
    \end{cases}
\end{equation*}

If the shading set \(D\) is non-empty we can modify definition of the directions
slightly
\begin{equation*}
    \begin{aligned}
        N &= \{(a,b+1),(a+1,b+1)\} \\
        E &= \{(a+1,b),(a+1,b+1)\} \\
        S &= \{(a,b),(a+1,b)\} \\
        W &= \{(a,b),(a,b+1)\}
    \end{aligned}
\end{equation*}
And we add the union of the sets in \(D\) into the mesh set \(R^\prime\).

\begin{definition}
    Given a pattern \(p\), define \(\addd{p}{(a,b)}\), and \\ \(\addi{p}{(a,b)}\),
    as the operations that return a mesh pattern equivalent to placing an
    decrease, or increase, in the center of box \((a,b)\), where \((a,b)\) is not
    in the mesh set of \(p\), in \(p\).
\end{definition}

\begin{example}
    \begin{align*}
        \addie = \pattern{}{1,2}{0/1,1/0,1/1,1/2,2/1}\\
        \addde = \pattern{}{2,1}{0/1,1/0,1/1,1/2,2/1}
    \end{align*}
\end{example}
We can define these formally as with \(\addp[D]{p}{(a,b)}\). Considering
only adding the ascent, adding a descent is very similar
\begin{equation*}
    \addi{p}{(a,b)} = (\pi^\prime, R^\prime)
\end{equation*}
\begin{equation*}
    \pi^\prime(i) = \begin{cases}
    \pi(i) & \text{if } i\neq a+1,a+2 \text{ and } \pi(i)<b\\
    \pi(i)+2 & \text{if } i\neq a+1,a+2 \text{ and } \pi(i)>b\\
    b+1, b+2 & \text{if } i = a+1, a+2
\end{cases}
\end{equation*}
and
\begin{equation*}
        R^\prime = \bigcup_{(i,j)\in R}r((i,j)) \cup 
\end{equation*}
Where \(r((i,j))\) is defined by
\begin{equation*}
    r((i,j)) =
    \begin{cases}
        \{(i,j)\} & \text{if } i<a, j<b\\
        \{(i,j),(i,j+1),(i,j+2)\} & \text{if } i<a, j=b\\
        \{(i,j+2)\} & \text{if } i<a, j>b\\
        \{(i,j),(i+1,j),(i+2,j)\} & \text{if } i=a, j<b\\
        \{(i,j+2),(i+1,j+2),(i+2,j+2)\} & \text{if } i=a, j>b\\
        \{(i+2,j)\} & \text{if } i>a, j<b\\
        \{(i+2,j),(i+2,j+1),(i+2,j+2)\} & \text{if } i>a, j=b\\
        \{(i+2,j+2)\} &\text{if } i>a, j>b\\
    \end{cases}
\end{equation*}

@@

It can be easily seen that in order to classify set equivalences one need only
consider coincidences within the family of mesh patterns with the same underlying
classical pattern, this is due to the fact that \(\perm{2,1}\in\av{\mperm{1,2}{R}}\)
and \(\perm{1,2}\in\av{\mperm{2,1}{R}}\) for all mesh-sets \(R\).

@@

By discussion of a number of rules we will show that the number of coincidence
classes follows the values shown in  \TableRef{tab:domclasses}. The experimental
data is calculated on permutations up to length 11.

\begin{table}[htb]
\begin{center}
\begin{tabularx}{\textwidth}{c|Y|Y|Y|Y|}
\cline{2-5}
& \multicolumn{4}{|c|}{Dominating Pattern}\\
\cline{2-5}
& \multicolumn{2}{|c|}{\(\perm{2,3,1}\)}&\multicolumn{2}{|c|}{\(\perm{3,2,1}\)}\\
\cline{2-5}
& \(\perm{1,2}\)&\(\perm{2,1}\)&\(\perm{1,2}\)&\(\perm{2,1}\)\\
\hline
\multicolumn{1}{|r|}{No Dominating rule}&\(220\)&\(220\)&\(220\)&\(220\)\\
\hline
\multicolumn{1}{|r|}{First Dominating rule}&\(85\)&\(43\)&\(220\)&\(29\)\\
\hline
\multicolumn{1}{|r|}{Second Dominating rule}&\(59\)&\(39\)&\(220\)&\(29\)\\
\hline
\multicolumn{1}{|r|}{Third Dominating rule}&\(56\)&\(39\)&\(220\)&\(29\)\\
\hline
\multicolumn{1}{|r|}{Experimental class size}&\(56\)&\(39\)&\(213\)&\(29\)\\
\hline
\end{tabularx}
\end{center}
    \caption{Class number reduction by application of Dominating Rules}
    \label{tab:domclasses}
\end{table}

\section{Coincidence classes of Av(\{321, (21, \textit{R})\}).}
Through experimentation we discover that there are at most 29 coincidence
classes of mesh patterns with underlying classical pattern \(\perm{2,1}\).

@@

    Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).
    % Consider a permutation \(\omega\) that contains an occurrence of \(m_2\)
    % by \NoteRef{not:downcmesh} any of these occurrences is also an occurence
    % of \(m_1\). This proves that every occurence of \(m_2\) is also an
    % occurence of \(m_1\) and therefore \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\).
    Since \(R_1\) is a subset of \(R_2\), \NoteRef{not:downcmesh} states that
    \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\)

    Now we consider a permutation \(\omega^\prime\in \av{\pi}\),
    suppose we have an occurrence of \(m_1\). Consider placing a point in the
    region corresponding to the box \((a,b)\), regardless of where in this
    region we place the point by condition~\ref{prop:dom1:cont} of the
    Proposition we create an occurrence of \(\pi\), therefore there can be no
    points in this region, which could have been represented in the mesh set
    \(R_1\) by adding the box \((a,b)\). Hence every occurrence of
    \(m_1\) is in fact an occurrence of \(m_2\), and we have that
    \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\).

@@

All coincidence classes of \(\av{\{\perm{3,2,1}, \mperm{2,1}{R}\}}\) can be
explained by repeated application of \PropositionRef{prop:dom1}.

This rule is understood very easily by seeing it in graphical form. In the
pattern in \FigureRef{fig:rule1} we can gain shading in three boxes since if
there is a point in any of these boxes we would gain an occurrence of the
dominating pattern \(\perm{3,2,1}\).

\begin{figure}[htb]
    \begin{center}
    \begin{tikzpicture}[scale=\picscale]
        \modpattern{}{2,1}{}
    \end{tikzpicture}
        \raisebox{2ex}{\(\mapsto\)}
    \begin{tikzpicture}[scale=\picscale]
        \filldraw (0.5,2.5) circle (3pt);
        \filldraw (1.5,1.5) circle (3pt);
        \filldraw (2.5,0.5) circle (3pt);
        \modpattern{}{2,1}{0/2,1/1,2/0}
    \end{tikzpicture}

@@

\section{Equivalence classes of Av(\{231, (21, \textit{R})\}).}
By application of \PropositionRef{prop:dom1} we obtain \(43\) equivalence
classes. Experimentation shows that there are in fact \(39\) equivalence
classes, for example the following two patterns are coincident in
\(\av{\perm{2,3,1}}\) but this is not explained by \PropositionRef{prop:dom1}.
\begin{equation*}
    m_1 = \pattern{}{2,1}{0/0,0/2,1/0,2/1} \text{ and } m_2 = \pattern{}{2,1}{0/0,0/2,1/0,1/1,2/1}
\end{equation*}

Consider an occurrence of \(m_1\) in a permutation in \(\av{\perm{2,3,1}}\), then if the region corresponding
to the box \((1,1)\) is empty we have an occurrence of \(m_2\). Otherwise, if there is any
increase in this box then we would have an occurrence of \(\perm{2,3,1}\),
however, since we are in \(\av{\perm{2,3,1}}\) this is not possible. This box
must therefore contain a decreasing subsequence. This gives rise to the following lemma:

@@

Going back to our example mesh patterns
\begin{equation*}
    \begin{tikzpicture}[scale=\picscale]
        \modpattern{}{2,1}{0/0,0/2,1/0,2/1}
        \filldraw (1.66,1.33) circle (3pt);
        \filldraw (1.33,1.66) circle (3pt);
    \end{tikzpicture}
\end{equation*}
We know that the region corresponding to the box \((1,1)\) contains a decreasing subsequence. The top point
of this decreasing subsequence can be chosen to act as the second point in the mesh pattern,
and therefore there are no points between the first point and the new second point.
Hence, we can shade this box as it is guaranteed to be empty. This shows that our
two example patterns are coincident.

This result generalises into the following rule for categorising equivalences
of mesh patterns in cases where there is a dominating classical pattern.

@@

    Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).

    Consider a permutation \(\omega\) that contains an occurrence of \(m_2\)
    by \NoteRef{not:downcmesh} any of these occurrences is also an occurence
    of \(m_1\). This proves that every occurence of \(m_2\) is also an
    occurence of \(m_1\) and therefore \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\).

    We will consider taking the first branch of every choice.
    Now consider a permutation in \(\omega^\prime \in \av{\pi}\). Suppose
    \(\omega^\prime\) contains \(m_1\) consider the region corresponding to
    \((a,b)\) in \(R_1\).

    If the region is empty, then we can freely shade the corresponding
    box \((a,b)\) in \(m_1\) and hence have an occurence of \(m_2\).

@@

    Similar arguments satisfy the remainder of the branches.

\end{proof}
\begin{figure}[ht]
\centering
\begin{tikzpicture}[scale = 0.75]

@@

By taking the First Dominating Pattern Rule and the Second Dominating Pattern
Rule together coincidences of classes of the form \(\av{\{\perm{2,3,1},\mperm{2,1}{R}\}}\)
are fully explained, obtaining 39 equivalence classes of mesh patterns.

\section{Equivalence classes of Av(\{231, (12, \textit{R})\}).}
When considering the equivalence classes of \(\av{\perm{2,3,1},\mperm{1,2}{R}}\)
we first apply the two Dominating Pattern rules previously established.
Starting from \(220\) classes, application of the first Dominating Pattern rule
gives \(85\) classes. Following this with the second Dominating Pattern rule
reduces the number of classes to \(59\). However we know that there are patterns
where the coincidences are not explained by the rules given above.

@@

Consider an occurrence of \(m_1\) in a permutation, if the region corresponding
to the box \((1,0)\) is empty then we have an occurrence of \(m_2\). Now look at
the case when this region is not empty. Consider choosing the rightmost point in
region.
% In order to rigorously establish this coincidence we need to consider what would
% happen if we were to choose a point in the box \((1,0)\) that we would like to shade.
% In order to get a chance of getting the shading we want consider choosing the
% rightmost point in the box.
This gives us the following mesh pattern.
\begin{equation*}
    \pattern{}{2,1,3}{0/3,2/0,2/1,3/0,3/1,3/2}
\end{equation*}
By application of the \PropositionRef{prop:dom1} we then achieve the following
mesh pattern
\begin{equation*}
    \pattern{2,3}{2,1,3}{0/3,1/3,2/0,2/1,3/0,3/1,3/2}
\end{equation*}
If we look at the highlighted points we see that the subpattern is an occurrence
of the mesh pattern that we originally desired. This gives rise to the following
rule:

\begin{proposition}[Third Dominating Pattern Rule]
    \label{prop:dom3}
    Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
    and a dominating classical pattern \(\pi = (\pi,\emptyset)\), the sets
    \(\av{\{\pi,m_1\}}\) and \(\av{\{\pi,m_2\}}\) are coincident if
    \begin{enumerate}
        \item \(R_1 \triangle R_2 = \{(a,b)\}\)
        \item\label{prop:dom3:condocc} \(\addp[D]{\mperm{\sigma}{R_1}}{(a,b)}\)
            is coincident with a mesh pattern containing an occurrence of
            \(\mperm{\sigma}{R_2}\).
    \end{enumerate}
\end{proposition}
\begin{proof}
    We need to prove that \(\av{\{\pi,m_1\}} = \av{\{\pi,m_2\}}\).

    Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).

    Consider a permutation \(\omega\) that contains an occurrence of \(m_2\)
    by \NoteRef{not:downcmesh} any of these occurrences is also an occurence
    of \(m_1\). This proves that every occurence of \(m_2\) is also an
    occurence of \(m_1\) and therefore \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\).

    Now consider a permutation \(\varrho\) in \(\av{\pi}\) that contains an
    occurrence of \(m_1\). If the region corresponding to the box \((a,b)\)
    is empty then we have an occurrence of \(m_2\). If the region is non-empty
    then by condition~\ref{prop:dom3:condocc}  of the proposition there exists
    a direction such that there exists an occurrence of a mesh pattern of
    length one longer than \(m_1\) in this position. This mesh pattern is
    coincident with another mesh pattern. This mesh pattern contains an
    occurrence of \(m_2\) so every occurrence of \(m_1\) is also an occurrence
    of \(m_2\).  Thus \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\) and the
    two patterns are coincident.
\end{proof}

@@

\section{Equivalence classes of Av(\{321, (12, \textit{R})\}).}
When considering equivalences of mesh patterns with underlying classical
pattern \(\perm{1,2}\) in \(\av{\perm{3,2,1}}\) application of the previously
established rules give no coincidences. Through experimentation we discover
that there are \(7\) equivalence classes which can be explained through the
use of two different lines of reasoning. Since the number of coincidences is
so small we will reason for these equivalences without attempting to generalise
into concrete rules.

>>>>

\chapter[Coincidence Classification]{Coincidences amongst families of mesh patterns and classical patterns}\label{chap:coincs}
One interesting question to ask about permutation patterns considers when a
pattern may be avoided by, or contained in, arbitrary permutations. Two
patterns \(\pi\) and \(\sigma\) are said to be \emph{coincident} if the set
of permutations that avoid \(\pi\) is the same as the set of permutations that
avoid \(\sigma\), \ie \(\av{\pi} = \av{\sigma}\). This extends to sets of patterns as well as single patterns.

We consider the avoidance sets, \(\av{\{\pi,q\}}\) where \(\pi\) is a classical pattern
of length \(3\) and \(q\) is a mesh pattern of length \(2\) in order to
establish some rules about when these two sets give the same avoidance set.
We fix \(\pi\) in order to define the  coincidence and say that \(\pi\) is the
\emph{dominating pattern}.

We first define some operations on mesh patterns.
\begin{definition}
Given a mesh pattern \(p\), let \(\addp[D]{p}{(a,b)}\) be the operation that returns
a mesh pattern equivalent to placing a point in the center of box \((a,b)\),
where \((a,b)\) is not in the mesh set of \(p\) , with shading defined by
\(D\subseteq\{N,E,S,W\}\).
The set \(D\) defines the shading by indicating that the boxes in the cardinal
directions in \(D\) next to the point are shaded in the resulting pattern.
\end{definition}
Since there is no ambiguity we let \(\addpe[D]\) be equivalent to \(\addp[D]{\varepsilon}{(0,0)}\).

\begin{example}
    The result of adding a single point to the empty permutation for each cardinal direction.
 \begin{align*}
     \addpe[\{N\}] = \textpattern{}{1}{0/1,1/1} &\qquad\addpe[\{E\}] =\textpattern{}{1}{1/1,1/0} \qquad \\
     \addpe[\{S\}] = \textpattern{}{1}{0/0,1/0} &\qquad\addpe[\{W\}] =\textpattern{}{1}{0/0,0/1} \qquad
\end{align*}
    A more complex example for \texttt{add\_point} could be
\begin{equation*}
\addp[\{N,E\}]{\pattern{}{2,1,3}{0/3}}{(2,3)} = \pattern{}{2,1,4,3}{0/3,0/4,2/4,3/4,3/3}
\end{equation*}
\end{example}
\begin{definition}
Formally, given a mesh pattern \(p = (\pi,R)\) \(\addp[D]{p}{(a,b)}\) gives a mesh
pattern \(p^\prime = (\pi^\prime,R^\prime)\) with length \(\setsize{\pi}+1\) defined by
\begin{equation*}
    \pi^\prime(i) = \begin{cases}
    \pi(i) & \text{if } i\neq a+1 \text{ and } \pi(i)<b\\
    \pi(i)+1 & \text{if } i\neq a+1 \text{ and } \pi(i)>b\\
    b+1 & \text{if } i = a+1
\end{cases}
\end{equation*}
and
\begin{equation*}
        R^\prime = \bigcup_{(i,j)\in R}r((i,j)),
\end{equation*}
where \(r((i,j))\) is defined by
\begin{equation*}
    r((i,j)) =
    \begin{cases}
        \{(i,j)\} & \text{if } i<a, j<b\\
        \{(i,j),(i,j+1)\} & \text{if } i<a, j=b\\
        \{(i,j+1)\} & \text{if } i<a, j>b\\
        \{(i,j),(i+1,j)\} & \text{if } i=a, j<b\\
        \{(i,j+1),(i+1,j+1)\} & \text{if } i=a, j>b\\
        \{(i+1,j)\} & \text{if } i>a, j<b\\
        \{(i+1,j),(i+1,j+1)\} & \text{if } i>a, j=b\\
        \{(i+1,j+1)\} &\text{if } i>a, j>b\\
    \end{cases}
\end{equation*}

If the shading set \(D\) is non-empty we can modify definition of the directions
slightly
\begin{equation*}
    \begin{aligned}
        N &= \{(a,b+1),(a+1,b+1)\} \\
        E &= \{(a+1,b),(a+1,b+1)\} \\
        S &= \{(a,b),(a+1,b)\} \\
        W &= \{(a,b),(a,b+1)\}
    \end{aligned}
\end{equation*}
And we add the union of the sets in \(D\) into the mesh set \(R^\prime\).
\end{definition}

\begin{definition}
    Given a pattern \(p\), define \(\addd{p}{(a,b)}\), and \\ \(\addi{p}{(a,b)}\),
    as the operations that return a mesh pattern equivalent to placing a
    decrease, or increase, in the center of box \((a,b)\), where \((a,b)\) is not
    in the mesh set of \(p\), in \(p\).
\end{definition}

\begin{example}
    \begin{align*}
        \addie = \pattern{}{1,2}{0/1,1/0,1/1,1/2,2/1}\\
        \addde = \pattern{}{2,1}{0/1,1/0,1/1,1/2,2/1}
    \end{align*}
\end{example}
We can define these formally as with \(\addp[D]{p}{(a,b)}\). Considering
only adding the ascent, as adding a descent is very similar
\begin{equation*}
    \addi{p}{(a,b)} = (\pi^\prime, R^\prime)
\end{equation*}
\begin{equation*}
    \pi^\prime(i) = \begin{cases}
    \pi(i) & \text{if } i\neq a+1,a+2 \text{ and } \pi(i)<b\\
    \pi(i)+2 & \text{if } i\neq a+1,a+2 \text{ and } \pi(i)>b\\
    b+j, b+j & \text{if } i = a+j, j\in\{1,2\}
\end{cases}
\end{equation*}
and
\begin{equation*}
        R^\prime = \bigcup_{(i,j)\in R}r((i,j)) \cup \text{crap to deal with strong ascent},
\end{equation*}
where \(r((i,j))\) is defined by
\begin{equation*}
    r((i,j)) =
    \begin{cases}
        \{(i,j)\} & \text{if } i<a, j<b\\
        \{(i,j),(i,j+1),(i,j+2)\} & \text{if } i<a, j=b\\
        \{(i,j+2)\} & \text{if } i<a, j>b\\
        \{(i,j),(i+1,j),(i+2,j)\} & \text{if } i=a, j<b\\
        \{(i,j+2),(i+1,j+2),(i+2,j+2)\} & \text{if } i=a, j>b\\
        \{(i+2,j)\} & \text{if } i>a, j<b\\
        \{(i+2,j),(i+2,j+1),(i+2,j+2)\} & \text{if } i>a, j=b\\
        \{(i+2,j+2)\} &\text{if } i>a, j>b\\
    \end{cases}
\end{equation*}

@@

It can be easily seen that in order to classify coincidences one need only
consider coincidences within the family of mesh patterns with the same underlying
classical pattern, this is due to the fact that \(\perm{2,1}\in\av{\mperm{1,2}{R}}\)
and \(\perm{1,2}\in\av{\mperm{2,1}{R}}\) for all mesh-sets \(R\).

@@

By discussion of a number of rules we will show that the number of coincidence
classes follows the values shown in  \TableRef{tab:domclasses}. The experimental
data in the last row of the table is calculated on permutations up to length 11.

\begin{table}[htb]
\begin{center}
\begin{tabularx}{\textwidth}{c|Y|Y|Y|Y|}
\cline{2-5}
& \multicolumn{4}{|c|}{Dominating Pattern}\\
\cline{2-5}
& \multicolumn{2}{|c|}{\(\perm{2,3,1}\)}&\multicolumn{2}{|c|}{\(\perm{3,2,1}\)}\\
\cline{2-5}
& \(\perm{1,2}\)&\(\perm{2,1}\)&\(\perm{1,2}\)&\(\perm{2,1}\)\\
\hline
\multicolumn{1}{|r|}{No Dominating rule}&\(220\)&\(220\)&\(220\)&\(220\)\\
\hline
\multicolumn{1}{|r|}{First Dominating rule}&\(85\)&\(43\)&\(220\)&\(29\)\\
\hline
\multicolumn{1}{|r|}{Second Dominating rule}&\(59\)&\(39\)&\(220\)&\(29\)\\
\hline
\multicolumn{1}{|r|}{Third Dominating rule}&\(56\)&\(39\)&\(220\)&\(29\)\\
\hline
\multicolumn{1}{|r|}{Experimental class size}&\(56\)&\(39\)&\(213\)&\(29\)\\
\hline
\end{tabularx}
\end{center}
    \caption{Coincidence class number reduction by application of Dominating Rules}
    \label{tab:domclasses}
\end{table}

\section{Coincidence classes of Av(\{321, (21, \textit{R})\}).}
Through experimentation, up to permutations of length 11, we discover that there
are at most 29 coincidence classes of mesh patterns with underlying classical
pattern \(\perm{2,1}\).

@@

    \noindent Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).
    % Consider a permutation \(\omega\) that contains an occurrence of \(m_2\)
    % by \NoteRef{not:downcmesh} any of these occurrences is also an occurence
    % of \(m_1\). This proves that every occurence of \(m_2\) is also an
    % occurence of \(m_1\) and therefore \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\).
    Since \(R_1\) is a subset of \(R_2\), \NoteRef{not:downcmesh} states that
    \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\)

    Now we consider a permutation \(\omega^\prime\in \av{\pi}\),
    containing an occurrence of \(m_1\). Consider placing a point in the
    region corresponding to the box \((a,b)\), regardless of where in this
    region we place the point by condition~\ref{prop:dom1:cont} of the
    Proposition we create an occurrence of \(\pi\), therefore there can be no
    points in this region, which could have been represented in the mesh set
    \(R_1\) by adding the box \((a,b)\). Hence every occurrence of
    \(m_1\) is in fact an occurrence of \(m_2\), and we have that
    \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\).

@@

All coincidence classes of \(\av{\{\perm{3,2,1}, \mperm{2,1}{R}\}}\) can be
explained by application of \PropositionRef{prop:dom1}. % repeated?

This rule can be understood very in graphical form. In the
pattern in \FigureRef{fig:rule1} we can gain shading in three boxes since if
there is a point in any of these boxes we would get an occurrence of the
dominating pattern \(\perm{3,2,1}\).

\begin{figure}[htb]
    \begin{center}
    \begin{tikzpicture}[scale=\picscale]
        \modpattern{}{2,1}{1/0,1/1,1/2,2/2}
    \end{tikzpicture}
        \raisebox{2ex}{\(\mapsto\)}
    \begin{tikzpicture}[scale=\picscale]
        \filldraw (2.5,0.5) circle (3pt);
        \modpattern{}{2,1}{1/0,1/1,1/2,2/2,2/0}
    \end{tikzpicture}

@@

Note that there are two natural extensions of this rule. We can replace \(\pi\)
with a set of classical patterns, or we can consider \(\pi\) to be a mesh pattern.

\section{Equivalence classes of Av(\{231, (21, \textit{R})\}).}
By application of \PropositionRef{prop:dom1} we obtain \(43\) coincidence
classes. Experimentation shows that there are in fact at most \(39\) coincidence
classes, for example the following two patterns are coincident in
\(\av{\perm{2,3,1}}\) but this is not explained by \PropositionRef{prop:dom1}.
\begin{equation*}
    m_1 = \pattern{}{2,1}{0/0,0/2,1/0,2/1} \text{ and } m_2 = \pattern{}{2,1}{0/0,0/2,1/0,1/1,2/1}
\end{equation*}

Consider an occurrence of \(m_1\) in a permutation in \(\av{\perm{2,3,1}}\), consisting
of elements \(x\) and \(y\). If the region corresponding
to the box \((1,1)\) is empty we have an occurrence of \(m_2\). Otherwise, if there is any
increase in this box then we would have an occurrence of \(\perm{2,3,1}\),
however, since we are in \(\av{\perm{2,3,1}}\) this is not possible. This box
must therefore contain a (non-empty) decreasing subsequence.
 This gives rise to the following lemma:

@@

Going back to our example mesh patterns
\begin{equation*}
    \begin{tikzpicture}[scale=\picscale]
        \modpattern{}{2,1}{0/0,0/2,1/0,2/1}
        \filldraw (1.66,1.33) circle (3pt);
        \filldraw (1.33,1.66) circle (3pt);
    \end{tikzpicture}
\end{equation*}
We know that the region corresponding to the box \((1,1)\) contains a decreasing subsequence.
If we let \(z\) be the topmost point in this decereasing subsequence, then \(xz\)
is an occurrence of \(m_2\). This shows that our two example patterns are coincident.

This result generalises into the following rule for categorising coincidences
of mesh patterns in cases where there is a dominating classical pattern.

@@

    \noindent Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).

    Consider a permutation \(\omega\) that contains an occurrence of \(m_2\).
    By \NoteRef{not:downcmesh} any of these occurrences is also an occurence
    of \(m_1\). This proves that every occurence of \(m_2\) is also an
    occurence of \(m_1\) and therefore \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\).

    We will consider taking the first branch of every choice.
    Now consider a permutation \(\omega^\prime \in \av{\pi}\). Suppose
    \(\omega^\prime\) contains \(m_1\) and consider the region corresponding to
    \((a,b)\) in \(R_1\).

    If the region is empty, the occurrence of \(m_1\) is trivially an occurence
    of \(m_2\).

@@

    Similar arguments satisfy the remainder of the branches.
\end{proof}
\begin{figure}[ht]
\centering
\begin{tikzpicture}[scale = 0.75]

@@

By taking the First Dominating Pattern Rule and the Second Dominating Pattern
Rule together coincidences of classes of the form \(\av{\{\perm{2,3,1},\mperm{2,1}{R}\}}\)
are fully explained, obtaining 39 coincidence classes of mesh patterns.

\section{Equivalence classes of Av(\{231, (12, \textit{R})\}).}
When considering the coincidence classes of \(\av{\perm{2,3,1},\mperm{1,2}{R}}\)
we first apply the two Dominating Pattern rules previously established.
Starting from \(220\) classes, application of the first Dominating Pattern rule
gives \(85\) classes. Following this with the second Dominating Pattern rule
reduces the number of classes to \(59\). However we know that there are patterns
where the coincidences are not explained by the rules given above.

@@

Consider an occurrence of \(m_1\) in a permutation, if the region corresponding
to the box \((1,0)\) is empty then we have an occurrence of \(m_2\). Now look at
the case when this region is not empty. Consider choosing the rightmost point in
region.
% In order to rigorously establish this coincidence we need to consider what would
% happen if we were to choose a point in the box \((1,0)\) that we would like to shade.
% In order to get a chance of getting the shading we want consider choosing the
% rightmost point in the box.
This gives us an occurrence of the following mesh pattern.
\begin{equation*}
    \pattern{}{2,1,3}{0/3,2/0,2/1,3/0,3/1,3/2}
\end{equation*}
By application of \PropositionRef{prop:dom1} we then achieve the following
mesh pattern
\begin{equation*}
    \pattern{2,3}{2,1,3}{0/3,1/3,2/0,2/1,3/0,3/1,3/2}
\end{equation*}
If we look at the highlighted points we see that the subpattern is an occurrence
of the mesh pattern that we originally desired. This gives rise to the following
rule:

\begin{proposition}[Third Dominating Pattern Rule]
    \label{prop:dom3}
    Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
    and a dominating classical pattern \(\pi = (\pi,\emptyset)\), the sets
    \(\av{\{\pi,m_1\}}\) and \(\av{\{\pi,m_2\}}\) are coincident if
    \begin{enumerate}
        \item \(R_1 \triangle R_2 = \{(a,b)\}\)
        \item\label{prop:dom3:condocc} \(\addp[D]{\mperm{\sigma}{R_1}}{(a,b)}\)
            where \(D\in\{N,E,S,W\}\)
            is coincident with a mesh pattern containing an occurrence of
            \(\mperm{\sigma}{R_2}\).
    \end{enumerate}
\end{proposition}
\begin{proof}
    We need to prove that \(\av{\{\pi,m_1\}} = \av{\{\pi,m_2\}}\).

    \noindent Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).

    Consider a permutation \(\omega\) that contains an occurrence of \(m_2\).
    By \NoteRef{not:downcmesh} any of these occurrences is also an occurence
    of \(m_1\). This proves that every occurence of \(m_2\) is also an
    occurence of \(m_1\) and therefore \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\).

    Now consider a permutation \(\varrho\) in \(\av{\pi}\) that contains an
    occurrence of \(m_1\). If the region corresponding to the box \((a,b)\)
    is empty then we have an occurrence of \(m_2\). If the region is non-empty
    then by condition~\ref{prop:dom3:condocc}  of the proposition there exists
    a direction such that there exists an occurrence of a mesh pattern of
    length one longer than \(m_1\) in this position. This mesh pattern is
    coincident with another mesh pattern that contains an
    occurrence of \(m_2\). Hence, every occurrence of \(m_1\) leads to an occurrence
    of \(m_2\).  Thus \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\) and the
    two patterns are coincident.
\end{proof}

@@

\section{Equivalence classes of Av(\{321, (12, \textit{R})\}).}
When considering coincidences of mesh patterns with underlying classical
pattern \(\perm{1,2}\) in \(\av{\perm{3,2,1}}\) application of the previously
established rules give no coincidences. Through experimentation we discover
that there are \(7\) non-trivial coincidence classes (all others are singletons)
which can be explained through the
use of two different lines of reasoning. Since the number of coincidences is
so small we will reason for these coincidences without attempting to generalise
into concrete rules.
